Automatic Matching of Bullet Lands

### 2016-01-12 imaging-paper.Rnw
fixed y label

Figure~\ref{fig:cms} shows the strong connection between the maximal number of consecutive striae and matches in the Hamby study. All 42 pairs of lands with at least thirteen CMS in common are matches. 
%
\begin{figure}[hbtp]
  \centering
%' \begin{subfigure}[b]{\textwidth}\centering
%' \caption{}
\begin{minipage}[t]{.47\textwidth}
<<cms-bars, echo=FALSE, fig.width=7, fig.height=3.5, out.width='\\textwidth'>>=
ggplot(data=subset(bstats, !flagged)) + 
  geom_bar(aes(x=factor(CMS))) + theme_bw() +
  theme(legend.position="bottom") + 
  xlab("maximum CMS")
@
%\end{subfigure}
%\begin{subfigure}[b]{\textwidth}\centering
\end{minipage}
\begin{minipage}[t]{.52\textwidth}
<<cms-spines, echo=FALSE,  fig.width=7, fig.height=4, out.width='\\textwidth'>>=
bstats$km <- c( "KNM", "KM")[as.numeric(bstats$match)+1]
ggplot(data=subset(bstats, !flagged)) + 
  geom_bar(aes(x=factor(CMS), fill=km), position="fill") +
  theme(legend.position="bottom") + 
  scale_fill_brewer("", palette="Paired") +
  xlab("maximum CMS")

@
\end{minipage}
%' \end{subfigure}
%' \begin{subfigure}[t]{\textwidth}\centering
%' \caption{\label{fig:cms-dens}Densities of maximum CMS for Known Matches (KM) and Known Non Matches (KNM).}
%' <<cms-densities, echo=FALSE,  fig.width=7, fig.height=3.5, out.width='.6\\textwidth'>>=
%' x <- seq(0,29, by=0.03)
%' km <- sm.density(bstats$CMS[bstats$match],h=2, model="Normal", display="none", eval.points=x)
%' knm <- sm.density(bstats$CMS[!bstats$match],h=1, display="none", eval.points=x)
%' dframe <- data.frame(x = x, km=km$estimate, knm=knm$estimate)
%' dframe$kmtail <- rev(cumsum(rev(dframe$km)))/sum(dframe$km)
%' dframe$knmtail <- rev(cumsum(rev(dframe$knm)))/sum(dframe$knm)
%' 
%' densities <- data.frame(
%'   match = rep(c("KM", "KNM"), each=(length(x)+2)),
%'   x = rep(c(0, x, 29), times = 2),
%'   estimate = c(0, km$estimate, 0, 0, knm$estimate,0))
%' 
%' ggplot() +
%'   geom_polygon(aes(x=x, y=estimate, fill=match, group=match), colour="black",
%'                data=densities, alpha=0.5) +
%' xlab("maximum CMS") +
%'   scale_fill_brewer("", palette="Paired") +
%'   ylab("Density estimates") +
%'   theme_bw() + 
%'   theme(legend.position=c(1,1), legend.justification=c(1,1),
%'         legend.background = element_rect(fill=alpha('white', 0.4))) 
%' #  geom_line(aes(x=x, y=pmin(0.2, kmtail/knmtail/10)), data=dframe) 
%' 
%' # ggplot(data=bstats) + 
%' #     geom_density(aes(x=cms, fill=match), alpha=0.5) +
%' #   xlab("maximum CMS") +
%' #   scale_fill_brewer("", palette="Paired") +
%' #   theme_bw() + 
%' #   theme(legend.position=c(1,1), legend.justification=c(1,1),
%' #         legend.background = element_rect(fill=alpha('white', 0.4))) 
%' 
%' @
%' \end{subfigure}
\caption{\label{fig:cms}Distribution of maximal CMS (left). Conditional barchart \citep{hummel} on the right: heights show probability of match/non-match given a specific CMS. All land-to-land comparisons with at least 13 CMS are matches.}
\end{figure}
%
There are two things that should be noted at this point: the automated algorithm finds a relatively high number of CMS even for non-matches. On average, there are \Sexpr{mean(bstats$CMS[!bstats$match], na.rm=T)} maximal CMS between known non-matches (with a standard deviation of \Sexpr{sd(bstats$CMS[!bstats$match])}). Known matches share on average \Sexpr{mean(bstats$CMS[bstats$match])} maximal CMS, with a standard deviation of \Sexpr{sd(bstats$CMS[bstats$match])}. While the probability for a match increases with the number of maximal CMS, a large number of maximal CMS by itself is not indicative of a match, as was previously pointed out by \citet{miller:1998}. Figure~\ref{fig:mismatch} shows a known mismatch, between two lands that share twelve consecutively matched striae. Visually we can easily tell that these two lands do not match well.

>>>>

Figure~\ref{fig:cms} shows the strong connection between the maximal number of consecutive striae and matches in the Hamby study. All 42 pairs of lands with at least thirteen CMS in common are matches. 
%
\begin{figure}[hbtp]
  \centering
%' \begin{subfigure}[b]{\textwidth}\centering
%' \caption{}
\begin{minipage}[t]{.47\textwidth}
<<cms-bars, echo=FALSE, fig.width=7, fig.height=3.5, out.width='\\textwidth'>>=
ggplot(data=subset(bstats, !flagged)) + 
  geom_bar(aes(x=factor(CMS))) + theme_bw() +
  theme(legend.position="bottom") + 
  xlab("maximum CMS")
@
%\end{subfigure}
%\begin{subfigure}[b]{\textwidth}\centering
\end{minipage}
\begin{minipage}[t]{.52\textwidth}
<<cms-spines, echo=FALSE,  fig.width=7, fig.height=4, out.width='\\textwidth'>>=
bstats$km <- c( "KNM", "KM")[as.numeric(bstats$match)+1]
ggplot(data=subset(bstats, !flagged)) + 
  geom_bar(aes(x=factor(CMS), fill=km), position="fill") +
  theme(legend.position="bottom") + 
  scale_fill_brewer("", palette="Paired") +
  xlab("maximum CMS") + ylab("Proportion")
@
\end{minipage}
%' \end{subfigure}
%' \begin{subfigure}[t]{\textwidth}\centering
%' \caption{\label{fig:cms-dens}Densities of maximum CMS for Known Matches (KM) and Known Non Matches (KNM).}
%' <<cms-densities, echo=FALSE,  fig.width=7, fig.height=3.5, out.width='.6\\textwidth'>>=
%' x <- seq(0,29, by=0.03)
%' km <- sm.density(bstats$CMS[bstats$match],h=2, model="Normal", display="none", eval.points=x)
%' knm <- sm.density(bstats$CMS[!bstats$match],h=1, display="none", eval.points=x)
%' dframe <- data.frame(x = x, km=km$estimate, knm=knm$estimate)
%' dframe$kmtail <- rev(cumsum(rev(dframe$km)))/sum(dframe$km)
%' dframe$knmtail <- rev(cumsum(rev(dframe$knm)))/sum(dframe$knm)
%' 
%' densities <- data.frame(
%'   match = rep(c("KM", "KNM"), each=(length(x)+2)),
%'   x = rep(c(0, x, 29), times = 2),
%'   estimate = c(0, km$estimate, 0, 0, knm$estimate,0))
%' 
%' ggplot() +
%'   geom_polygon(aes(x=x, y=estimate, fill=match, group=match), colour="black",
%'                data=densities, alpha=0.5) +
%' xlab("maximum CMS") +
%'   scale_fill_brewer("", palette="Paired") +
%'   ylab("Density estimates") +
%'   theme_bw() + 
%'   theme(legend.position=c(1,1), legend.justification=c(1,1),
%'         legend.background = element_rect(fill=alpha('white', 0.4))) 
%' #  geom_line(aes(x=x, y=pmin(0.2, kmtail/knmtail/10)), data=dframe) 
%' 
%' # ggplot(data=bstats) + 
%' #     geom_density(aes(x=cms, fill=match), alpha=0.5) +
%' #   xlab("maximum CMS") +
%' #   scale_fill_brewer("", palette="Paired") +
%' #   theme_bw() + 
%' #   theme(legend.position=c(1,1), legend.justification=c(1,1),
%' #         legend.background = element_rect(fill=alpha('white', 0.4))) 
%' 
%' @
%' \end{subfigure}
\caption{\label{fig:cms}Distribution of maximal CMS (left). Conditional barchart \citep{hummel} on the right: heights show probability of match/non-match given a specific CMS. All land-to-land comparisons with at least 13 CMS are matches.}
\end{figure}
%
There are two things that should be noted at this point: the automated algorithm finds a relatively high number of CMS even for non-matches. On average, there are \Sexpr{mean(bstats$CMS[!bstats$match], na.rm=T)} maximal CMS between known non-matches (with a standard deviation of \Sexpr{sd(bstats$CMS[!bstats$match])}). Known matches share on average \Sexpr{mean(bstats$CMS[bstats$match])} maximal CMS, with a standard deviation of \Sexpr{sd(bstats$CMS[bstats$match])}. While the probability for a match increases with the number of maximal CMS, a large number of maximal CMS by itself is not indicative of a match, as was previously pointed out by \citet{miller:1998}. Figure~\ref{fig:mismatch} shows a known mismatch, between two lands that share twelve consecutively matched striae. Visually we can easily tell that these two lands do not match well.
